---
title: "Assignment 5-Decision tree and randomForest"
author: "Andrew Abisha Hu, Elaine Su, Eric Xiong, Tianye Wang"
date: "9/6/2018"
output: html_document
---

Here is the description of variables:

```
category: the category to which the project belongs
goal: the project’s raising goal ($)
pledged: $ pledged at the end of the fund-raising period
funded: “TRUE” if the goal is achieved (i.e., pledged >= goal); “FALSE” otherwise
blurb_adj: ratio of counts of adjectives to total word counts of the blurb
blurb_noun: ratio of counts of nouns to total word counts of the blurb
blurb_verb: ratio of counts of verbs to total word counts of the blurb
blurb_total: total word counts of the blurb on the project page
full_adj: ratio of counts of adjectives to total word counts of the main content
full_noun: ratio of counts of nouns to total word counts of the main content
full_verb: ratio of counts of verbs to total word counts of the main content
full_total: total word counts in the main content the project page
train: = 1 if the project is in the training set and 0 otherwise
```

1. Use the training data set to create a regression tree and predict pledged by only
full_total. Use mindev=1e-6 to overgrow the tree. Plot the partition of the tree model.
(provide R codes, 0.5%)

```{r setup, include=FALSE}
setwd("/Users/andrewhu/Documents/GitHub/Machine-Learning/Class Practice/Assignment 5- Decision tree and randomForest")

library(tree)
ks <- read.csv("KickStarter2.csv")
kick <- read.csv("KickStarter2.csv")


attach(ks)
fit= tree(pledged~full_total,data=kick[kick$train==1,],mindev=1e-6)

partition.tree(fit, main="Fitted Function")
```


2. Prune the tree in Question (1) by requesting the number of terminal nodes to be 8. Plot
the decision tree and the partition of the pruned tree model.
```{r }
fit2 = prune.tree(fit, best=8)
par(mfrow=c(1,2))
plot(fit2, type= "uniform")
text(fit2, cex=.7)
partition.tree(fit2, cex=.8)
```

3. According to the pruned tree, for a project with content in 1800 words, what is the
predicted pledged amount? How many projects in the data also have the same predicted
value? (0.5%)

```{r}
predict(fit2, data.frame(full_total=1800))
sum(ifelse(full_total==1800,1,0))
```

4. Use the training data set to create a regression tree to predict pledged by all variables
except funded. Use mindev=1e-6 to overgrow the tree and then prune it by requesting the
number of terminal nodes to be 10. Plot the decision tree and label the decision tree plot.
What are the key predictors used to split this tree?

```{r}
fit3= tree(pledged~category+goal+blurb_adj+blurb_noun+blurb_verb+blurb_total+full_adj+full_noun+full_verb+full_total,data=kick[kick$train==1,],mindev=1e-6)


fit4 = prune.tree(fit3,best=10)

plot(fit4, type="uniform")
text(fit4)
```

```
full_total, goal, category, blurb_verb, full_noun
```

5. Further prune the tree you pruned in question (4) by requesting six terminal nodes.
Which predictor is no longer a key predictor after you re-prune the tree?

```{r}
fit5 = prune.tree(fit4,best=6)
plot(fit5, type= "uniform")
text(fit5)
```


```
full_noun
```

6. Use the testing set to compute MSE = mean((pledged – pledged_hat)2) for the two
pruned tree models in Questions (4) and (5). Which one performs better in terms of test
MSE? (provide R codes, 0.5%)

```{r}
yhat1<- predict(fit4,ks[train==0,])
yhat2<- predict(fit5,ks[train==0,])

mean((ks$pledged[!train]-yhat1)^2, na.rm=T)
mean((ks$pledged[!train]-yhat2)^2, na.rm=T) #better
```

7. According to the better tree in Question (6), what is the predicted pledged amount for
an art project with a goal of $10000 if the campaign runner writes 10 words in the blurb
with 4 verbs and 1000 words in the main content with 124 nouns? (0.5%)

```{r}
predict(fit5, data.frame(category="Art",goal=10000,blurb_total=10,blurb_verb=0.4,full_total=1000,full_noun=0.124,blurb_adj=0,blurb_noun=0,full_adj=0,full_verb=0))
```

8. Use the training data set to run “bagging” by using the randomForest() function and
predict pledged by all variables except funded. Create a chart to show the importance of
predictors in terms of %IncMSE. What are the three most important predictors? Use
set.seed(12345) to specify the random seed. (provide R codes, 1%)

```{r}
library(randomForest)
set.seed(12345)

fit8= randomForest(pledged~category+goal+blurb_adj+blurb_noun+blurb_verb+blurb_total+full_adj+full_noun+full_verb+full_total,data=kick[kick$train==1,],mtry=10,importance=T)

varImpPlot(fit8)
```

9. Now use the training data set to run “random forest” by using the randomForest()
function and predict pledged by all variables except funded. Create a chart to show the
importance of predictors in terms of %IncMSE. What are the three most important
predictors? Use set.seed(12345) to specify the random seed. 

```{r}
set.seed(12345)

fit9= randomForest(pledged~category+goal+blurb_adj+blurb_noun+blurb_verb+blurb_total+full_adj+full_noun+full_verb+full_total,data=kick[kick$train==1,],importance=T)

varImpPlot(fit9)
```


10. Compute test MSE for the bagging and random forest models in Questions (8) and (9).
Which one performs better in terms of test MSE?

```{r}
yhat3<- predict(fit8,ks[train==0,])
yhat4<- predict(fit9,ks[train==0,])

mean((kick$pledged[train==0]-yhat3)^2, na.rm=T)
mean((kick$pledged[train==0]-yhat4)^2, na.rm=T)
#randomForest is better
```

11. Create a partial dependence plot to depict the marginal effect of full_total using the
random forest model with the test data.

```{r}
fit= randomForest(pledged~category+goal+blurb_adj+blurb_noun+blurb_verb+blurb_total+full_adj+full_noun+full_verb+full_total,data=kick[kick$train==1,],importance=T)

partialPlot(fit, kick[kick$train==0,], full_total)
```

12. Use the training data and run a “random forest” to predict funded by all variables
except pledged. Create a chart to show the importance of predictors in terms of
MeanDecreaseAccuracy. What are the three most important predictors? Use
set.seed(12345) to specify the random seed. 

```{r}
fit= randomForest(as.factor(funded)~category+goal+blurb_adj+blurb_noun+blurb_verb+blurb_total+full_adj+full_noun+full_verb+full_total,data=kick[kick$train==1,],importance=T)

varImpPlot(fit)
```

13. Based on your findings, would you conclude that full_total has a causal effect on
funded? Why? 

```
Yes, because it is the most important predictor of funded according to MeanDecreaseAccuracy
in the random forest model.
```

---

FordUS.csv records the number of vehicles sold under the brand “Ford” in the US from
July 2011 to February 2018. (data from
http://shareholder.ford.com/financials/monthly-sales-reports).

```
Month: the calendar month
US_sales: the number of vehicles sold under the brand “Ford” in the US in the
month
```

14. Read the data into R and use ts() to make US_sales a time-series object in R. Create a
plot of US_sales over time. 

```{r}
ford<- read.csv("FordUS.csv")
attach(ford)
ts= ts(US_sales, start=c(2011,7), frequency = 12)
plot(ts, type="b")
```

15. Based on the trajectory of US_sales over time, comment on the trend and seasonality
of the sales. Which model may better describe the data, an additive model or a
multiplicative model? Why?

```
Since the amplitude of seasonality seems constant, an additive model may better describe the
data.
```

16. Use the data from July 2011 to June 2017 as your training data and the rest as your
test data. Fit the training data with an additive Halt-Winters model. What are the values of
the smoothing parameters: alpha, beta, and gamma? (provide R codes, 0.5%) Hint: use
window() to obtain a subset of a time-series object


```{r}
train= window(ts, start=c(2011,7), end=c(2017,6))
fit9 =HoltWinters(train)
fit9

```